---
title: "Project proposal"
author: "Piping Hot"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(here)
```

## 1. Introduction

Do factors such as age, hypertension and BMI increase the risk of a stroke?

We got our dataset from the Kaggle website. The data was collected confidentially because it concerns health so the details about its collection have not been published. Each case is a different individual. The 12 variables refer to the attributes of the individuals.

## 2. Data

```{r}
stroke_risk <- read.csv(here::here("data/healthcare-dataset-stroke-data.csv"))

stroke_risk <- stroke_risk %>%
  mutate(stroke = if_else(stroke == "1", "Yes", "No"))

glimpse(stroke_risk)
```

## 3. Data analysis plan

We expect to see a higher proportion of stroke with those of a higher age bracket, higher bmi, higher glucose level, have hypertension, private work type, people who have smoked, and urban.

```{r}
stroke_risk %>%
  filter(smoking_status != "Unknown") %>%
  count(smoking_status, stroke) %>%
  ggplot(aes(x = stroke, y = n, fill = smoking_status)) +
  geom_col(position = "fill") +
  scale_fill_viridis_d() +
  labs(x = "Stroke", y = "Proportion", fill = "Smoking status") +
  theme_minimal()

stroke_risk %>%
  count(gender, stroke) %>%
  filter(gender != "Other")

```

Comparison groups; gender, age, hypertension, heart_disease, ever_married, work_type, Residence_type, avg_glucose_level, bmi, smoking_status.

Conditional probability (helps us calculate proportions), mean (helps us to compare continuous variables for those who have and haven't had a stroke), IQR (helps us to visualise the spread of  the data)

