---
title: "Project proposal"
author: "Piping Hot"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(here)
library(ggridges)
```

## 1. Introduction

Do factors such as age, hypertension and BMI increase the risk of a stroke?

We got our dataset from the Kaggle website. The data was collected confidentially because it concerns health so the details about its collection have not been published. Each case is a different individual. The 12 variables refer to the attributes of the individuals.

## 2. Data

```{r}
stroke_risk <- read.csv(here::here("data/healthcare-dataset-stroke-data.csv"))

stroke_risk <- stroke_risk %>%
  mutate(stroke = if_else(stroke == "1", "Yes", "No"))

glimpse(stroke_risk)
```

## 3. Data analysis plan

Is stroke risk affected by 

```{r}
stroke_risk %>%
  filter(smoking_status != "Unknown") %>%
  count(smoking_status, stroke) %>%
  ggplot(aes(x = stroke, y = n, fill = smoking_status)) +
  geom_col(position = "fill") +
  scale_fill_viridis_d()

stroke_risk %>%
  count(gender, stroke) %>%
  filter(gender != "Other")
```

The outcome (response, Y) and predictor (explanatory, X) variables you will use to answer your question.
The comparison groups you will use, if applicable.
The statistical method(s) that you believe will be useful in answering your question(s). (You can update these later as you work on your project.)
What results from these specific statistical methods are needed to support your hypothesized answer?

