---
title: "Modelling of Strokes (temp)"
subtitle: "With respect to health and lifestyle"
author: "Pipping Hot <br> Ishan, Anwen and Alex"
institute: "University of Edinburgh"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(xaringanthemer)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, include=FALSE}
stroke_risk <- read.csv(here::here("data/healthcare-dataset-stroke-data.csv"))
stroke_risk <- stroke_risk %>%
  mutate(
    #hypertension = if_else(hypertension == 1, TRUE, FALSE),
    #heart_disease = if_else(heart_disease == 1, TRUE, FALSE),
    #stroke = if_else(stroke == 1, TRUE, FALSE),
    bmi = if_else(bmi == "N/A", NA_real_, as.numeric(bmi))
  )
```

```{r include=FALSE, echo = FALSE}
style_xaringan(
  title_slide_background_image = "none"
)
```

class: center, middle

## Do factors such as age, hypertension and BMI increase the risk of a stroke?

---

class: inverse, center, middle

# Introductory plots
---

# Plot 1

```{r plot-1, echo = FALSE}
stroke_risk %>%
  mutate(stroke = if_else(stroke == 1, "Yes", "No"), stroke = fct_rev(stroke)) %>%
  ggplot(aes(x = avg_glucose_level, fill = stroke)) +
  geom_density(alpha = 0.5) +
  labs(
    x = "Average Glucose Level",
    y = "Density",
    title = "Average glucose level compared to stroke risk",
    fill = "Stroke"
  ) +
  scale_fill_viridis_d() +
  theme_minimal()
```

---

# Plot 2

```{r plot-2, echo = FALSE}
median_age_stroke <- stroke_risk %>%
  mutate(stroke = if_else(stroke == "1", "Yes", "No"), stroke = fct_rev(stroke)) %>%
  group_by(stroke) %>%
  summarise(median_age = median(age))

median_age_stroke %>%
  ggplot(aes(stroke, median_age, fill = stroke)) +
  geom_col() +
  scale_fill_manual(values = c("Purple", "Orange")) +
  labs(
    x = "Stroke", 
    y = "Midean Age", 
    title = "The Average Age of Patients who have had a Stroke",
    fill = "Stroke"
  ) +
  theme(legend.position = "none") +
  theme_minimal()
```

class: inverse, middle, center

# Using xaringan

---

# xaringan

- The presentation is created using the `xaringan` package

- Use `---` to separate slides and `--` for incremental builds

--

- Like this

---

# Layouts

You can use plain text

- or bullet points

.pull-left[
or text in two columns $^*$
]
.pull-right[
- like
- this
]

.footnote[
[*] And add footnotes
]

---

# Code

```{r boring-regression}
# a boring regression
model <- lm(dist ~ speed, data = cars)
tidy(model)
glance(model)
```

---

# Plots

```{r recode-species, echo = FALSE}
# In this chunk I'm doing a bunch of analysis that I don't want to present 
# in my slides. But I need the resulting data frame for a plot I want to present.
iris_modified <- iris %>%
  mutate(Species = fct_other(Species, keep = "setosa"))
```

```{r plot-iris, echo = FALSE}
# Code hidden with echo = FALSE
# Uses modified iris dataset from previous chunk
# Play around with height and width until you're happy with the look
ggplot(data = iris_modified, mapping = aes(x = Sepal.Width, y = Sepal.Length, color = Species)) +
  geom_point() + 
  theme_minimal() # theme options: https://ggplot2.tidyverse.org/reference/ggtheme.html
```

---

## Plot and text

.pull-left[
- Some text
- goes here
]
.pull-right[
```{r warning=FALSE, out.width="100%", fig.width=4, echo=FALSE}
# see how I changed out.width and fig.width from defaults
# to make the figure bigger
ggplot(penguins, aes(x = bill_length_mm, y = species, color = species)) +
  geom_boxplot() +
  theme_minimal()
```
]

---

# Tables

If you want to generate a table, make sure it is in the HTML format (instead of Markdown or other formats), e.g.,

```{r iris-table, echo = FALSE}
kable(head(iris), format = "html")
```

---

# Images

```{r castle, echo = FALSE, out.width = "60%", fig.align = "center", fig.cap = "Image credit: Photo by Jörg Angeli on Unsplash."}
include_graphics("https://images.unsplash.com/photo-1535448033526-c0e85c9e6968?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1650&q=80")
```

Or you can also include a full page image. See next slide.

---

background-image: url(https://images.unsplash.com/photo-1535448033526-c0e85c9e6968?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1650&q=80)

---

# Math Expressions

You can write LaTeX math expressions inside a pair of dollar signs, e.g. &#36;\alpha+\beta$ renders $\alpha+\beta$. You can use the display style with double dollar signs:

```
$$\bar{X}=\frac{1}{n}\sum_{i=1}^nX_i$$
```

$$\bar{X}=\frac{1}{n}\sum_{i=1}^nX_i$$

Limitations:

1. The source code of a LaTeX math expression must be in one line, unless it is inside a pair of double dollar signs, in which case the starting `$$` must appear in the very beginning of a line, followed immediately by a non-space character, and the ending `$$` must be at the end of a line, led by a non-space character;

1. There should not be spaces after the opening `$` or before the closing `$`.

1. Math does not work on the title slide (see [#61](https://github.com/yihui/xaringan/issues/61) for a workaround).

---

# Feeling adventurous?

- Want to find out more about `xaringan`? See https://slides.yihui.name/xaringan/#1.

- You are welcomed to use the default styling of the slides. In fact, that's what I expect majority of you will do. You will differentiate yourself with the content of your presentation.

- But some of you might want to play around with slide styling. The 
`xaringanthemer` provides some solutions for this that: https://pkg.garrickadenbuie.com/xaringanthemer.

- And if you want more bells and whistles, there is also `xaringanExtra`: https://pkg.garrickadenbuie.com/xaringanExtra.
